---
title: 'Lab #3'
author: "Amy Crawford"
date: "February 16, 2017"
output: pdf_document
---

```{r, warning = F, message = F}
library(tidyverse)
library(readxl)

#setwd("~/Documents/STAT 585X/In Class Activities/Week6/Lab3")

f1_s1 <- readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=1) 
f1_s2 <- readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=2)
f1_s3 <- readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=3)
f1_s4 <- readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=4)

f1_s1$semester <- "Sem1"
f1_s2$semester <- "Sem2"
f1_s3$semester <- "Sem3"
f1_s4$semester <- "Sem4"

f2_s1 <- readxl::read_excel("Spreadsheets/FileTwo.xlsx", sheet=1)
f2_s2 <- readxl::read_excel("Spreadsheets/FileTwo.xlsx", sheet=2)
f2_s3 <- readxl::read_excel("Spreadsheets/FileTwo.xlsx", sheet=3)
f2_s4 <- readxl::read_excel("Spreadsheets/FileTwo.xlsx", sheet=4)

f2_s1$semester <- "Sem1"
f2_s2$semester <- "Sem2"
f2_s3$semester <- "Sem3"
f2_s4$semester <- "Sem4"
```


## Looking at the data
It seems that all four sheets in file 1 share column names with the exception of sheet 2. This sheet has an extra column that consists completely of `NAs`. We remove this column and check whether columns names match.
```{r}
f1_s2[,15] <- NULL
names(f1_s1) == names(f1_s2)
```

Now we will stack all four of the sheets for file #1 to make one large dataset.
```{r}
file1 <- rbind(f1_s1, f1_s2, f1_s3, f1_s4)
file1$test <- "Test1"
names(file1) <- c("ID", "Pre_Post", names(file1[,3:20]))
```

Taking a peek at file 2, we see that sheet 1 has one fewer column than the rest of the other 3 sheets.....
```{r}
f2_s1 <- cbind(f2_s1[,1:16], `MTH 3` = rep(NA, times = nrow(f2_s1)), f2_s1[,17:ncol(f2_s1)])
names(f2_s1) <- names(f2_s2)
file2 <- rbind(f2_s1, f2_s2, f2_s3, f2_s4)

file2$test <- "Test2"
names(file2) <- c("ID", "Pre_Post", names(file2[,3:42]))

```


## Preparing the data sets to be merged

File 1
```{r}
#Split by pre_post
file1_pre <- file1 %>% filter(Pre_Post %in% c("Pre", "PRE")) %>% as.data.frame()
file1_pre_clean <- file1_pre %>% gather(key = Answer, value = Value, 3:13)

file1_post <- file1 %>% filter(Pre_Post %in% c("Post", "POST")) %>% as.data.frame()
file1_post_clean <- file1_post %>% gather(key = Answer, value = Value, 3:13)
file1_post_clean <- file1_post_clean[,-c(3:7)]



file1_clean <- full_join(file1_pre_clean, file1_post_clean, by = c("ID", "semester","test", "Pre_Post", "Answer", "Value")) %>% as.data.frame() 
```

Copy basic information from "Pre" to "Post"
```{r}
#Change PRE to Pre, change POST to post
#Copy GENDER, Characteristics, Treatment Part1 and Treatment Part 2 from "Pre" to "Post" of same ID.

file1_clean <- file1_clean %>% mutate(Pre_Post = replace(Pre_Post, Pre_Post=="PRE", "Pre"))
file1_clean <- file1_clean %>% mutate(Pre_Post = replace(Pre_Post, Pre_Post=="POST", "Post"))
file1_clean <- file1_clean %>% mutate(
  `Normalized Changes` = replace(`Normalized Changes`, 
                                 Pre_Post=="Post",file1_clean$`Normalized Changes`[Pre_Post=="Pre"]))
file1_clean <- file1_clean %>% mutate(
  GENDER = replace(GENDER, Pre_Post=="Post",file1_clean$GENDER[Pre_Post=="Pre"]))
file1_clean <- file1_clean %>% mutate(
  Characteristic= replace( Characteristic, Pre_Post=="Post",file1_clean$ Characteristic[Pre_Post=="Pre"]))
file1_clean <- file1_clean %>% mutate(
  `Treatment Part 1`= replace(`Treatment Part 1`,                               Pre_Post=="Post",file1_clean$ `Treatment Part 1`[Pre_Post=="Pre"]))
file1_clean <- file1_clean %>% mutate(
  `Treatment Part 2 `= replace(`Treatment Part 2 `, Pre_Post=="Post",
                               file1_clean$ `Treatment Part 2 `[Pre_Post=="Pre"]))

file1_clean %>%  arrange(ID) %>% head()  ##clean file1
```

File2
```{r}
file2_pre <- file2 %>% filter(Pre_Post %in% c("Pre", "PRE")) %>% as.data.frame()
file2_pre_clean <- file2_pre %>% gather(key = Answer, value = Value, 3:35)

file2_post <- file2 %>% filter(Pre_Post %in% c("Post", "POST")) %>% as.data.frame()
file2_post_clean <- file2_post %>% gather(key = Answer, value = Value, 3:35)
file2_post_clean <- file2_post_clean[,-c(3:7)]



file2_clean <- full_join(file2_pre_clean, file2_post_clean, by = c("ID", "semester","test", "Pre_Post", "Answer", "Value")) %>% as.data.frame() 

```

Copy basic information from "Pre" to "Post"
```{r}
#Change PRE to Pre, change POST to post
#Copy GENDER, Characteristics, Treatment Part1 and Treatment Part 2 from "Pre" to "Post" of same ID.

file2_clean <- file2_clean %>% mutate(Pre_Post = replace(Pre_Post, Pre_Post=="PRE", "Pre"))
file2_clean <- file2_clean %>% mutate(Pre_Post = replace(Pre_Post, Pre_Post=="POST", "Post"))
file2_clean <- file2_clean %>% mutate(
  `Normalized Change` = replace(`Normalized Change`, 
                                 Pre_Post=="Post",file2_clean$`Normalized Change`[Pre_Post=="Pre"]))
file2_clean <- file2_clean %>% mutate(
  GENDER = replace(GENDER, Pre_Post=="Post",file2_clean$GENDER[Pre_Post=="Pre"]))
file2_clean <- file2_clean %>% mutate(
  Characteristic= replace( Characteristic, Pre_Post=="Post",file2_clean$ Characteristic[Pre_Post=="Pre"]))
file2_clean <- file2_clean %>% mutate(
  `Treatment Part 1`= replace(`Treatment Part 1`,Pre_Post=="Post",
                              file2_clean$ `Treatment Part 1`[Pre_Post=="Pre"]))
file2_clean <- file2_clean %>% mutate(
  `Treatment Part 2`= replace(`Treatment Part 2`, Pre_Post=="Post",
                               file2_clean$ `Treatment Part 2`[Pre_Post=="Pre"]))

file2_clean %>% arrange(ID) %>% head()  ##clean file2
```

Make column name of semester 1 and semester 2 the same
```{r}
colnames(file1_clean) <- colnames(file2_clean)
```

Summarize total value and normalized value of semester 1 and 2
```{r}
file1_sum <- file1_clean %>% filter(Answer  == "Total")
file2_sum <- file2_clean %>% filter(Answer  == "Total")
file_sum <- rbind(file1_sum, file2_sum)  ##clean data summary
```


