---
title: 'Lab #3'
author: "Amy Crawford"
date: "February 16, 2017"
output: pdf_document
---

```{r, warning = F, message = F}
library(tidyverse)
library(readxl)
setwd("~/Documents/STAT 585X/In Class Activities/Week6/Lab3")

f1_s1 <- readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=1) 
f1_s2 <- readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=2)
f1_s3 <- readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=3)
f1_s4 <- readxl::read_excel("Spreadsheets/FileOne.xlsx", sheet=4)

f1_s1$semester <- "Sem1"
f1_s2$semester <- "Sem2"
f1_s3$semester <- "Sem3"
f1_s4$semester <- "Sem4"

f2_s1 <- readxl::read_excel("Spreadsheets/FileTwo.xlsx", sheet=1)
f2_s2 <- readxl::read_excel("Spreadsheets/FileTwo.xlsx", sheet=2)
f2_s3 <- readxl::read_excel("Spreadsheets/FileTwo.xlsx", sheet=3)
f2_s4 <- readxl::read_excel("Spreadsheets/FileTwo.xlsx", sheet=4)

f2_s1$semester <- "Sem1"
f2_s2$semester <- "Sem2"
f2_s3$semester <- "Sem3"
f2_s4$semester <- "Sem4"
```


## Looking at the data
It seems that all four sheets in file 1 share column names with the exception of sheet 2. This sheet has an extra column that consists completely of `NAs`. We remove this column and check whether columns names match.
```{r}
f1_s2[,15] <- NULL
names(f1_s1) == names(f1_s2)
```

Now we will stack all four of the sheets for file #1 to make one large dataset.
```{r}
file1 <- rbind(f1_s1, f1_s2, f1_s3, f1_s4)
file1$test <- "Test1"
names(file1) <- c("ID", "Pre_Post", names(file1[,3:20]))
```

Taking a peek at file 2, we see that sheet 1 has one fewer column than the rest of the other 3 sheets.....
```{r}
f2_s1 <- cbind(f2_s1[,1:16], `MTH 3` = rep(NA, times = nrow(f2_s1)), f2_s1[,17:ncol(f2_s1)])
names(f2_s1) <- names(f2_s2)
file2 <- rbind(f2_s1, f2_s2, f2_s3, f2_s4)

file2$test <- "Test2"
names(file2) <- c("ID", "Pre_Post", names(file2[,3:42]))

```


## Preparing the data sets to be merged

File 1
```{r}
#Split by pre_post
file1_pre <- file1 %>% filter(Pre_Post %in% c("Pre", "PRE")) %>% as.data.frame()
file1_pre_clean <- file1_pre %>% gather(key = Answer, value = Value, 3:13)

file1_post <- file1 %>% filter(Pre_Post %in% c("Post", "POST")) %>% as.data.frame()
file1_post_clean <- file1_post %>% gather(key = Answer, value = Value, 3:13)
file1_post_clean <- file1_post_clean[,-c(3:7)]



file1_clean <- full_join(file1_pre_clean, file1_post_clean, by = c("ID", "semester","test", "Pre_Post", "Answer", "Value")) %>% as.data.frame() 
```

File2
```{r}



```





